---
title: "MEP-LINCs Pilot Analysis"
date: "`r Sys.Date()`"
output: pdf_document
---



```{r setup, echo=FALSE, message=FALSE}
library("ggplot2")
library("data.table")
library("MEMA")
library("grid")

numericMedian <- function(x) as.numeric(median(x))

evalMedians <- function(values, reps){
  tmp <- median(values[reps], na.rm=TRUE)
}

create8WellPseudoImage <- function(DT, pr, prDisplay){
  p <- ggplot(DT, aes_string(x="ArrayColumn", y="ArrayRow",colour=pr))+
    geom_point(size=1)+
    scale_y_reverse()+   scale_x_continuous(breaks= c(min(DT$ArrayColumn),round(mean(c(min(DT$ArrayColumn),max(DT$ArrayColumn)))),max(DT$ArrayColumn)))+
    scale_colour_gradient(low = "white", high = "red",limits=quantile(DT[[pr]],probs = c(0,.998),na.rm=TRUE))+
    guides(colour = guide_legend(prDisplay, keywidth = .5, keyheight = .5))+
    ggtitle(paste(prDisplay,"for",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+
    xlab("")+ylab("")+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=4)
}

create8WellHistograms <- function(DT, pr, prDisplay, binwidth = diff(quantile(DT[[pr]],probs = c(0,.98),na.rm=TRUE))/50, upperProb = .99, ncol = 4) {
  p <- ggplot(DT, aes_string(x=pr))+
    geom_histogram(binwidth = binwidth)+
    scale_x_continuous(limits = quantile(DT[[pr]],probs = c(0,upperProb),na.rm=TRUE))+
    ggtitle(paste(prDisplay,"in",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+
    xlab(prDisplay)+
        theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=ncol)
}

#Set the staining set to be analyzed
# ss <- "SS1"
# cmp <- "Cyto."
ss <- "SS2"
cmp <- ""

popDT <- fread(paste0("./",ss,"/Population/Annotated Data/PC3_",ss,"_PopAnn.txt"))

cDT <- fread(paste0("./",ss,"/Cell/Annotated Data/PC3_",ss,"_CellAnn.txt"))

sDT <- fread(paste0("./",ss,"/Cell/Annotated Data/PC3_",ss,"_CellSpotAnn.txt"))

barcodes <- sort(unique(cDT$Barcode))

#Set a threshold for filtering wells on their QA score
wellQAThresh <- 0

#TODO: Read this from sDT
lthresh <- 0.6

# Clean up the dataset that will be used in the analysis
popDT <- popDT[!popDT$ShortName =="blank"]
popDT$ShortName <- gsub("blank","",popDT$ShortName)
setnames(popDT,"Ligand","LigandAnnotID")
ms <- gregexpr("^[[:alnum:]]*[^_]", popDT$LigandAnnotID)
popDT$Ligand <- unlist(regmatches(popDT$LigandAnnotID, ms))


```
##Summary
This experiment uses the MEP-LINCs 
`r ss` staining set on `r unique(sDT$CellLine)` cells in 8 well MEMAs. 

##Introduction
The plates were printed with 35 row by 20 column MEMAs using a 4x7 pin head that printed 5x5 blocks. Each sample spot contains one ECM protein paired with Collagen I. There are 46 different ECM proteins in the array arranged in a random fashion.

Images of each well were gathered on a Tecan LS Reloaded laser scanner and Olympus ScanR automated microscope. This staining set includes, DAPI, `r unique(sDT$Endpoint488)` (488nm), `r unique(sDT$Endpoint555)` (532 and 555nm) and `r unique(sDT$Endpoint647)` (635 and 647nm). Data from DAPI staining is only gathered by the ScanR. 

Tecan data is gathered at the spot population level by fitting round regions of interest (ROIs) to each spot. The Tecan data in this report uses the net values defined as the raw ROI value minus the mean of the local background.

The ScanR data comes from the nucleus as defined by the DAPI staining and a cytoplasmic annulus of 10 pixels that surrounds the nucleus.

The spots that were not printed are labeled as blank.  

\newpage

##QA Scoring of the dataset
Each well is scored for even cell seeding according the count of the DAPI-stained nuclei.

```{r Heatmaps_QAScores, echo=FALSE, fig.width=3.7,fig.height=4}

for (barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ShortName)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "SpotCellCount",prDisplay = "Spot Cell Count")
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = "LoessSCC",prDisplay = "Loess Spot Cell Count")
    suppressWarnings(print(p))
    
  p <- create8WellHistograms(DT,pr = "SpotCellCount", prDisplay = "Spot Cell Count")
      suppressWarnings(print(p))

  
  wellScores <- unique(DT[,list(Well, QAScore=sprintf("%.2f",QAScore))])

  p <- ggplot(DT, aes(x=LoessSCC))+
    geom_histogram(binwidth=.04)+
    geom_vline(xintercept=lthresh, colour="blue")+
    geom_text(data=wellScores, aes(label=paste0("QA\n",QAScore)), x = 1.5, y = 40, size = rel(3), colour="red")+
    ggtitle(paste("QA on Loess Model of Spot Cell Count for",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Spot Cell Count")+xlim(0,3)+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=4)
  suppressWarnings(print(p))
    
  cat(sprintf("Mean QA Score for %9s = %.2f \n",barcode,mean(DT$QAScore)))

}

```

\newpage

##Stain Pseudoimages
The pseudoimages of each well's raw signals are shown in the plots below.

```{r Pseudoimages_all_stains, echo=FALSE, fig.width=3.7,fig.height=4}

for (barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ShortName)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "Mean.Intensity.DAPI", prDisplay = "Mean DAPI")
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.488"), prDisplay = unique(DT$Endpoint488))
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.555") ,prDisplay = unique(DT$Endpoint555))
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.647") ,prDisplay = unique(DT$Endpoint647))
  suppressWarnings(print(p))
  
}

```

\newpage

##Z Height Pseudoimages
The pseudoimages of the focus Z heights are shown in the plots below.

```{r Z_Height_Pseudoimages, echo=FALSE, fig.width=3.7,fig.height=4}

for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ShortName)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "Z",prDisplay = "Z Height")
  suppressWarnings(print(p))
  
}

for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ShortName)
  DT <- DT[!"fiducial"]
  
  p <- ggplot(DT, aes(x=factor(Well), y=Z))+
    geom_boxplot()+
    ggtitle(paste("Z Height By Well for Plate", barcode))+
    xlab("Well")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
  suppressWarnings(print(p))
  
}

```

```{r Z_vs_ECM,echo=FALSE, fig.width=8, fig.height=4}
for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ShortName)
  DT <- DT[!"fiducial"]
  
p <- ggplot(DT, aes(x=reorder(ShortName, Z, FUN=median), y=Z))+
  geom_boxplot()+
  ggtitle(paste("Z Height By ECM for Plate",barcode))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))
}

```

\newpage

##Univariate Signal Distributions
The next step in the EDA is to look at the raw univariate cell-level signal responses.

```{r Univariate Signals, echo=FALSE, fig.width=4}

for (barcode in barcodes){
  setkey(cDT, Barcode)
  DT <- cDT[barcode]
  
  p <- create8WellHistograms(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.488"), prDisplay = unique(DT$Endpoint488))
  suppressWarnings(print(p))
 
    p <- create8WellHistograms(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.555"), prDisplay = unique(DT$Endpoint555))
  suppressWarnings(print(p))
  
    p <- create8WellHistograms(DT, pr = paste0(cmp,"Mean.Intensity.Alexa.647"), prDisplay = unique(DT$Endpoint647), binwidth = .5, upperProb = .97)
  suppressWarnings(print(p))
  
    p <- create8WellHistograms(DT, pr = "Density", prDisplay = "Density", ncol = 3)
  suppressWarnings(print(p))
  
}
```


##Filtering
Wells with QA scores below `r wellQAThresh ` are removed from further analysis due to quality issues in the cell level data. The HighSerum control wells are also excluded from the plots.


```{r Filtering, echo=FALSE}

passedQAWells <- unique(sDT[sDT$QAScore > wellQAThresh,list(Barcode,Well)])


setkey(cDT,Barcode,Well)
cDT <- merge(cDT,passedQAWells)

setkey(sDT,Barcode,Well)
sDT <- merge(sDT,passedQAWells)

setkey(cDT,Ligand)
cDT <- cDT[!"HighSerum"]

setkey(sDT,Ligand)
sDT <- sDT[!"HighSerum"]

```


```{r Bivariate Signals, echo=FALSE,eval=FALSE, fig.width=4}

for (barcode in barcodes){
  setkey(cDT, Barcode)
  DT <- cDT[barcode]

  p <- ggplot(DT, aes(x=Cyto.Mean.Intensity.Alexa.488.MedNorm, y=Cyto.Mean.Intensity.Alexa.647.MedNorm, colour=ShortName))+
    geom_point(size=.5,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint647),"vs", unique(DT$Endpoint488),"in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab(unique(DT$Endpoint488))+ylab(unique(DT$Endpoint647))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
     p <- ggplot(DT, aes(x=SpotCellCount, y=Cyto.Mean.Intensity.Alexa.488.MedNorm, colour=ShortName))+
    geom_point(size=.5,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint488),"vs", "Spot Cell Count","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Spot Cell Count")+ylab(unique(DT$Endpoint488))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
       p <- ggplot(DT, aes(x=SpotCellCount, y=Cyto.Mean.Intensity.Alexa.647.MedNorm, colour=ShortName))+
    geom_point(size=.5,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint647),"vs", "Spot Cell Count","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Spot Cell Count")+ylab(unique(DT$Endpoint647))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
}
```


```{r Bivariate_density, echo=FALSE,, fig.width=4, eval=FALSE}

for (barcode in barcodes){
  setkey(sDT, Barcode)
  DT <- sDT[barcode]

    p <- ggplot(DT, aes(x=Density, y=SpotCellCount, colour=ShortName))+
    geom_point(size=1,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste("SpotCellCount","vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab("SpotCellCount")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
   p <- ggplot(DT, aes(x=Density, y=Mean.Intensity.DAPI, colour=ShortName))+
    geom_point(size=2,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste("Mean DAPI","vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab("Mean DAPI")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
   p <- ggplot(DT, aes(x=Density, y=Cyto.Mean.Intensity.Alexa.647.MedNorm, colour=ShortName))+
    geom_point(size=2,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint647),"vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab(unique(DT$Endpoint647))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
   p <- ggplot(DT, aes(x=Density, y=Cyto.Mean.Intensity.Alexa.488.MedNorm, colour=ShortName))+
    geom_point(size=2,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint488),"vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab(unique(DT$Endpoint488))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
  p <- ggplot(DT, aes(x=Density, y=Cyto.Mean.Intensity.Alexa.555.MedNorm, colour=ShortName))+
    geom_point(size=2,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint555),"vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab(unique(DT$Endpoint555))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
 
   
   p <- ggplot(DT, aes(x=Density, y=Cyto.Mean.Intensity.Alexa.647.MedNorm, colour=ShortName))+
    geom_point(size=2,alpha=.5)+
    guides(colour=guide_legend(ncol = 2,override.aes = list(alpha = 1,size=2)))+
    ggtitle(paste(unique(DT$Endpoint647),"vs", "Density","in", unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Density")+ylab(unique(DT$Endpoint647))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)),legend.key.size = unit(.3, "cm"))+
    facet_wrap(~Ligand, ncol=4)
  suppressWarnings(print(p))
  
   
}
```