---
output: pdf_document
---
```{r, echo=FALSE}
#Set the staining set to be analyzed (SS1|SS2|SS3)
 ss <- "SS2"
```
---
title: "MEP-LINCs `r ss` Pilot Analysis"
date: "`r Sys.Date()`"
output: pdf_document
---



```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library("ggplot2")
library("data.table")
library("MEMA")
library("grid")


create8WellPseudoImage <- function(DT, pr, prDisplay){
  p <- ggplot(DT, aes_string(x="ArrayColumn", y="ArrayRow",colour=pr))+
    geom_point(size=1)+
    scale_y_reverse()+   scale_x_continuous(breaks= c(min(DT$ArrayColumn),round(mean(c(min(DT$ArrayColumn),max(DT$ArrayColumn)))),max(DT$ArrayColumn)))+
    scale_colour_gradient(low = "white", high = "red",limits=quantile(DT[[pr]],probs = c(0,.998),na.rm=TRUE))+
    guides(colour = guide_legend(prDisplay, keywidth = .5, keyheight = .5))+
    ggtitle(paste(prDisplay,"for",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+
    xlab("")+ylab("")+theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1),plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=4)
}

create8WellHistograms <- function(DT, pr, prDisplay, binwidth = diff(quantile(DT[[pr]],probs = c(0,.98),na.rm=TRUE))/50, upperProb = .99, ncol = 4) {
  p <- ggplot(DT, aes_string(x=pr))+
    geom_histogram(binwidth = binwidth)+
    scale_x_continuous(limits = quantile(DT[[pr]],probs = c(0,upperProb),na.rm=TRUE))+
    ggtitle(paste(prDisplay,"in",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+
    xlab(prDisplay)+
        theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=ncol)
}

cDT <- fread(paste0("./",ss,"/Cell/Annotated Data/PC3_",ss,"_CellAnn.txt"), showProgress = FALSE)

sDT <- fread(paste0("./",ss,"/Cell/Annotated Data/PC3_",ss,"_CellSpotAnn.txt"))

mepDT <- fread(paste0("./",ss,"/Cell/Annotated Data/PC3_",ss,"_CellMepAnn.txt"))

barcodes <- sort(unique(cDT$Barcode))

#Set a threshold for filtering wells on their QA score
wellQAThresh <- 0

#TODO: Read this from sDT
lthresh <- 0.6



```
##Summary
This experiment uses the MEP-LINCs 
`r ss` staining set on `r unique(sDT$CellLine)` cells in 8 well MEMAs. 

##Introduction
The plates were printed with 35 row by 20 column MEMAs using a 4x7 pin head that printed 5x5 blocks. Each sample spot contains one ECM protein paired with Collagen I. There are 46 different ECM proteins in the array arranged in a random fashion.

Images of each well were gathered on a Tecan LS Reloaded laser scanner and Olympus ScanR automated microscope. This staining set includes, DAPI, `r unique(sDT$Endpoint488)` (488nm), `r unique(sDT$Endpoint555)` (532 and 555nm) and `r unique(sDT$Endpoint647)` (635 and 647nm). Data from DAPI staining is only gathered by the ScanR. 

Tecan data is gathered at the spot population level by fitting round regions of interest (ROIs) to each spot. The Tecan data in this report uses the net values defined as the raw ROI value minus the mean of the local background.

The ScanR data comes from the nucleus as defined by the DAPI staining and a cytoplasmic annulus of 10 pixels that surrounds the nucleus.

The spots that were not printed are labeled as blank.  

\newpage

##QA Scoring of the dataset
Each well is scored for even cell seeding according the count of the DAPI-stained nuclei.

```{r Heatmaps_QAScores, echo=FALSE, fig.width=3.7,fig.height=4, eval=TRUE}

for (barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "SpotCellCount",prDisplay = "Spot Cell Count")
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = "LoessSCC",prDisplay = "Loess Spot Cell Count")
    suppressWarnings(print(p))
    
  p <- create8WellHistograms(DT,pr = "SpotCellCount", prDisplay = "Spot Cell Count")
      suppressWarnings(print(p))

  
  wellScores <- unique(DT[,list(Well, QAScore=sprintf("%.2f",QAScore))])

  p <- ggplot(DT, aes(x=LoessSCC))+
    geom_histogram(binwidth=.04)+
    geom_vline(xintercept=lthresh, colour="blue")+
    geom_text(data=wellScores, aes(label=paste0("QA\n",QAScore)), x = 1.5, y = 40, size = rel(3), colour="red")+
    ggtitle(paste("QA on Loess Model of Spot Cell Count for",unique(DT$CellLine), "cells in plate",unique(DT$Barcode)))+xlab("Spot Cell Count")+xlim(0,3)+
    theme(axis.text.x = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(.5)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))+
    facet_wrap(~Well, ncol=4)
  suppressWarnings(print(p))
    
  cat(sprintf("Mean QA Score for %9s = %.2f \n",barcode,mean(DT$QAScore)))

}

```

\newpage

##Stain Pseudoimages
The pseudoimages of each well's raw signals are shown in the plots below.

```{r Pseudoimages_all_stains, echo=FALSE, fig.width=3.7,fig.height=4}

for (barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "MeanIntensityDAPI", prDisplay = "Mean DAPI")
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = "MeanIntensityAlexa488", prDisplay = unique(DT$Endpoint488))
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = "MeanIntensityAlexa555",prDisplay = unique(DT$Endpoint555))
  suppressWarnings(print(p))
  
  p <- create8WellPseudoImage(DT, pr = "MeanIntensityAlexa647",prDisplay = unique(DT$Endpoint647))
  suppressWarnings(print(p))
  
}

```

\newpage

##Z Height Pseudoimages
The pseudoimages of the focus Z heights are shown in the plots below.

```{r Z_Height_Pseudoimages, echo=FALSE, fig.width=3.7,fig.height=4, eval=FALSE}

for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  
  p <- create8WellPseudoImage(DT, pr = "Z",prDisplay = "Z Height")
  suppressWarnings(print(p))
  
}

for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  
  p <- ggplot(DT, aes(x=factor(Well), y=Z))+
    geom_boxplot()+
    ggtitle(paste("Z Height By Well for Plate", barcode))+
    xlab("Well")+ylab("")+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.5)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
  suppressWarnings(print(p))
  
}

```

```{r Z_vs_ECM,echo=FALSE, fig.width=8, fig.height=4, eval=FALSE}
for(barcode in barcodes){
  DT <-sDT[sDT$Barcode==barcode,]
  #Remove the fiducial entries
  setkey(DT,ECMp)
  DT <- DT[!"fiducial"]
  
p <- ggplot(DT, aes(x=reorder(ECMp, Z, FUN=median), y=Z))+
  geom_boxplot()+
  ggtitle(paste("Z Height By ECM for Plate",barcode))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))
}

```

\newpage

##Univariate Signal Distributions
The next step in the EDA is to look at the raw univariate cell-level signal responses.

```{r Univariate Signals, echo=FALSE, fig.width=4, eval=FALSE}

for (barcode in barcodes){
  setkey(cDT, Barcode)
  DT <- cDT[barcode]
  
  p <- create8WellHistograms(DT, pr = "MeanIntensityAlexa488", prDisplay = unique(DT$Endpoint488))
  suppressWarnings(print(p))
 
    p <- create8WellHistograms(DT, pr = "MeanIntensityAlexa555", prDisplay = unique(DT$Endpoint555))
  suppressWarnings(print(p))
  
    p <- create8WellHistograms(DT, pr = "MeanIntensityAlexa647", prDisplay = unique(DT$Endpoint647), binwidth = .5, upperProb = .97)
  suppressWarnings(print(p))
  
    p <- create8WellHistograms(DT, pr = "Density", prDisplay = "Density", ncol = 3)
  suppressWarnings(print(p))
  
}
```

\newpage

##Cell Cycle Plots


```{r Cell_cycle_plots,echo=FALSE, fig.width=8, fig.height=5, eval=FALSE}

    for (barcode in unique(cDT$Barcode)){
  mDT <- cDT[cDT$Barcode == barcode]
  p <- ggplot(mDT, aes(x=TotalIntensityDAPI))+geom_bar(binwidth = 50000)+
  facet_wrap(~Well, nrow=2)+
    xlim(0,quantile(mDT$TotalIntensityDAPI,probs=.98, na.rm=TRUE))+
    ggtitle(paste("Total DAPI Signal"))+
    ylab("Count")+xlab("Total Intensity DAPI")+
    theme(strip.text = element_text(size = 5), axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  suppressWarnings(print(p))
    }
  
if (ss == "SS1"){
 
  
} else if (ss == "SS2"){

for (barcode in unique(cDT$Barcode)){
  mDT <- cDT[cDT$Barcode == barcode]
  p <- ggplot(mDT, aes(x=TotalIntensityDAPI,y=log2(MeanIntensityAlexa647), colour=EdUPositive))+geom_point(size = .4)+
      geom_rug(col=rgb(.5,0,0,alpha=.01))+
  facet_wrap(~Well, nrow=2)+
    xlim(0,quantile(mDT$TotalIntensityDAPI,probs=.98, na.rm=TRUE))+
    ggtitle(paste("Cell Cycle",unique(mDT$Endpoint647), "Signal"))+
    ylab(paste("Well Normalized",unique(mDT$Endpoint647)))+xlab("")+
    theme(strip.text = element_text(size = 5), axis.text.x = element_blank(), axis.ticks.x = element_blank())
  suppressWarnings(print(p))
}
}
```


##Filtering
Wells with QA scores below `r wellQAThresh ` are removed from further analysis due to quality issues in the cell level data. The HighSerum control wells are also excluded from the plots.


```{r Filtering, echo=FALSE}

passedQAWells <- unique(sDT[sDT$QAScore > wellQAThresh,list(Barcode,Well)])

setkey(cDT,Barcode,Well)
cDT <- merge(cDT,passedQAWells)

setkey(cDT,Ligand)
cDT <- cDT[!"HighSerum"]

setkey(cDT,ECMp)
cDT <- cDT[!"fiducial"]

cDT <- cDT[!cDT$SpotCellCount < 40]

setkey(sDT,Barcode,Well)
sDT <- merge(sDT,passedQAWells)

setkey(sDT,Ligand)
sDT <- sDT[!"HighSerum"]

#Filtering of MEP dataset
setkey(mepDT,ECMp)
mepDT <-mepDT[!"fiducial"]
mepDT <-mepDT[!"blank"]

setkey(mepDT,Ligand)
mepDT <- mepDT[!"HighSerum"]

```

\newpage

##MEP Perimeter Box Plots

```{r Perimeter_Median_Ratios, echo=FALSE,fig.width=8, fig.height=4, eval=FALSE}

calcGroupRatios <- function(x,group,signal){
  #browser()
  medianInGroup <- median(x[[signal]][x[[group]]], na.rm=TRUE)
  medianOutGroup <- median(x[[signal]][!x[[group]]], na.rm=TRUE)
  return(medianInGroup/medianOutGroup)
} 

if(ss=="SS1"){
  
cDT <- cDT[, MitoPerimeterRatio := calcGroupRatios(.SD,group="Perimeter",signal="MeanIntensityAlexa647MedNorm"), by="Ligand,ECMp"]

cDT <- cDT[, MitoSparseRatio := calcGroupRatios(.SD,group="Sparse",signal="MeanIntensityAlexa647MedNorm"), by="Ligand,ECMp"]

setkey(cDT,MitoPerimeterRatio)
MEP <- unique(cDT[,list(ECMp,Ligand,MitoPerimeterRatio,MitoSparseRatio,Barcode,Well,ArrayRow,ArrayColumn,SpotCellCount,MeanIntensityAlexa647SpotNorm)])

p <- ggplot(cDT, aes(x=factor(Perimeter), y=log2(MeanIntensityAlexa647)))+geom_boxplot(notch=TRUE)+
      xlab("Perimeter")+ylab(paste(unique(cDT$Endpoint647), "Mean Intensity (log2)"))+
    ggtitle(paste(unique(cDT$Endpoint647), "Mean Intensity in Interior vs. Perimeter Cells"))
  suppressWarnings(print(p))
  
  p <- ggplot(cDT, aes(x=factor(Sparse), y=log2(MeanIntensityAlexa647)))+geom_boxplot(notch=FALSE)+
    xlab("Sparse")+ylab(paste(unique(cDT$Endpoint647), "Mean Intensity (log2)"))+
    ggtitle(paste(unique(cDT$Endpoint647), "Mean Intensity in Dense vs. Sparse Cells"))
  suppressWarnings(print(p))
  
  p <- ggplot(cDT, aes_string(x="factor(ECMp)",y=paste0("MitoPerimeterRatio"), colour="Ligand"))+geom_point(size=3)+
    ggtitle(paste("Perimeter vs Interior Ratios in ",unique(cDT$Endpoint647), "Signal by MEP"))+
    ylab(paste(unique(cDT$Endpoint647),"Perimeter to Interior Ratio"))+xlab("")+
    theme(strip.text = element_text(size = 5), axis.text.x = element_text(angle = 90),
          axis.ticks.x = element_blank())
  suppressWarnings(print(p))
  
} else if (ss=="SS2"){

cDT <- cDT[, EdUPerimeterRatio := calcGroupRatios(.SD,group="Perimeter",signal="MeanIntensityAlexa647MedNorm"), by="Ligand,ECMp"]

cDT <- cDT[, EdUSparseRatio := calcGroupRatios(.SD,group="Sparse",signal="MeanIntensityAlexa647MedNorm"), by="Ligand,ECMp"]

#TODO debug here and get to a MEP level dataset
setkey(cDT,EdUPerimeterRatio)
MEP <- unique(cDT[,list(ECMp,Ligand,EdUPerimeterRatio,EdUSparseRatio,Barcode,Well, ArrayRow,ArrayColumn,SpotCellCount,EdUPositiveProportion)])

p <- ggplot(cDT, aes(x=factor(Perimeter), y=log2(MeanIntensityAlexa647SpotNorm)))+geom_boxplot(notch=FALSE)+
  ylim(0,1)+
  facet_wrap(~ECMp)
  suppressWarnings(print(p))
  
  p <- ggplot(cDT, aes(x=factor(Perimeter), y=log2(MeanIntensityAlexa647SpotNorm)))+geom_boxplot(notch=FALSE)+
  ylim(0,1)+
  facet_wrap(~Ligand)
  suppressWarnings(print(p))
  
  p <- ggplot(cDT, aes(x=factor(Sparse), y=log2(MeanIntensityAlexa647SpotNorm)))+geom_boxplot(notch=FALSE)+
  ylim(0,1)+
  facet_wrap(~ECMp)
  suppressWarnings(print(p))
  
  p <- ggplot(cDT, aes(x=factor(Sparse), y=log2(MeanIntensityAlexa647SpotNorm)))+geom_boxplot(notch=FALSE)+
  ylim(0,1)+
  facet_wrap(~Ligand)
  suppressWarnings(print(p))
  
  
    p <- ggplot(MEP, aes(x=EdUSparseRatio))+geom_histogram(binwidth=.01)
  #ylim(0,1)+
  #facet_wrap(~Ligand)
  suppressWarnings(print(p))

} else if(ss == "SS3"){
  
} else stop(paste(ss, "is an invlalid value for the ss staining set parameter"))
```

```{r EdU_Perimeter_1st, echo=FALSE,fig.width=8, fig.height=11, eval=FALSE}
nrLigands <- length(unique(cDT$Ligand))
for(ligand in unique(cDT$Ligand)[1:floor(nrLigands/2)]){
  mDT <- cDT[cDT$Ligand == ligand]
  p <- ggplot(mDT, aes(x=1,y=log2(MeanIntensityAlexa647SpotNorm), colour=Perimeter))+geom_boxplot(notch=FALSE)+
  facet_wrap(~ECMp, nrow=3)+
    ylim(-4,4)+
    ggtitle(paste("Perimeter vs Interior",unique(mDT$Endpoint647), "Signal for",ligand, "Paired MEPs"))+
    ylab(paste("Spot Normalized",unique(mDT$Endpoint647)))+xlab("")+
    theme(strip.text = element_text(size = 5), axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  suppressWarnings(print(p))
}
```

```{r EdU_Perimeter_2nd, echo=FALSE,fig.width=8, fig.height=10, eval=FALSE}

for(ligand in unique(cDT$Ligand)[floor(nrLigands/2):nrLigands]){
  mDT <- cDT[cDT$Ligand == ligand]
  p <- ggplot(mDT, aes(x=1,y=log2(MeanIntensityAlexa647SpotNorm), colour=Perimeter))+geom_boxplot(notch=FALSE)+
  facet_wrap(~ECMp, nrow=3)+
    ylim(-4,4)+
    ggtitle(paste("Perimeter vs Interior",unique(mDT$Endpoint647), "Signal for",ligand, "Paired MEPs"))+
    ylab(paste("Spot Normalized",unique(mDT$Endpoint647)))+xlab("")+
    theme(strip.text = element_text(size = 5), axis.text.x = element_blank(),
          axis.ticks.x = element_blank())
  suppressWarnings(print(p))
}
```

\newpage

##H3 Response

The following plots look at the plate normalized responses of the `r unique(mepDT$Endpoint488)` signal stratified by ligand, ECM protein and MEP.

```{r 488_response, echo=FALSE}

p <- ggplot(mepDT, aes(x = reorder(Ligand, MeanIntensityAlexa488MedNorm, FUN=median), y = MeanIntensityAlexa488MedNorm))+geom_boxplot()+
  ggtitle(paste(unique(mepDT$Endpoint488),"Plate Normalized Ligand Response for",unique(mepDT$CellLine), "Cells"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

p <- ggplot(mepDT, aes(x = reorder(ECMpAnnotID, MeanIntensityAlexa488MedNorm, FUN=median), y = MeanIntensityAlexa488MedNorm))+geom_boxplot()+
  ggtitle(paste(unique(mepDT$Endpoint488),"Plate Normalized ECM Protein Response for",unique(mepDT$CellLine), "Cells"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

p <- ggplot(mepDT, aes(x = MeanIntensityAlexa488MedNorm ))+geom_histogram(binwidth = .03)+
  ggtitle(paste(unique(mepDT$Endpoint488),"Plate Normalized MEP  Response for",unique(mepDT$CellLine), "Cells"))+
  xlab("")+ylab("")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(.8)), axis.text.y = element_text(angle = 0, vjust = 0.5, hjust=1, size=rel(1)), plot.title = element_text(size = rel(1)),legend.text=element_text(size = rel(.3)),legend.title=element_text(size = rel(.3)))
suppressWarnings(print(p))

setkey(mepDT, MeanIntensityAlexa488MedNorm)

```

The following lists are the bottom and top 25 MEPS ordered by their normalized `r unique(mepDT$Endpoint488)` values. These responses were divided by the high serum responses, then the high serum responses were filtered from the dataset.

  
```{r, echo=FALSE}
setkey(mepDT, MeanIntensityAlexa488MedNorm)
print(mepDT[1:25,list(Ligand,ECMp,MeanIntensityAlexa488MedNorm)])

print(mepDT[(nrow(mepDT)-25):nrow(mepDT),list(Ligand,ECMp,MeanIntensityAlexa488MedNorm)])

```
